Trends from 30-Year Observations of Downward Solar Irradiance in Thessaloniki, Greece

### 2023-12-28 MDPI_submission.Rmd
Commit 2023-12-28 19:09

Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
For time series with a uniform trend, we would expect
the CUSUMs of the anomalies to have a symmetric 'V' shape centered
around the middle of the series. This would indicate that the anomalies
are evenly distributed around the climatological mean, and for a
positive uniform trend, the first half is below and the second half
above the climatological mean.
In our case, there is a more complex
evolution of the anomalies.
For all skies (Figure\ \ref{fig:cusummonth-1}),
we observe three
rather distinct periods: a) a downward part between the start of the
datasets and about 2000, denoting that all anomalies are negative, thus
below the climatology; b) a relatively steady part lasting for almost 20
years suggesting little variability in SDR anomalies; and c) a steep
upward part to the present indicating anomalies above the climatology. The
CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behavior with some short-term
differences that do not change the overall pattern.
For clear skies
(Figure\ \ref{fig:cusummonth-2}), a monotonic downward tendency is evident until 2004,
suggesting that the anomalies are all negative. After 2004 the anomalies
turn to positive at a fast rate for about five years and at a slower
rate thereafter.

@@

In order to unveil further the features of the variability of the three
datasets, Figure\ \ref{fig:cusumnotrendmonthly}
presents another set of CUSUM plots using anomalies
after the long-term linear trend is removed. With this approach, periods
when the CUSUMs diverge from zero can be interpreted as a systematic
variation of SDR from the climatological mean. When the CUSUM is
increasing, the anomalies values are above the climatology and vice
versa. 
Overall, for all- and cloudy-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-3}) we
observe periods with anomalies diverging from the climatological values,
each lasting for several years. These fluctuations are probably within
the natural variability and no distinct changes are identified. The
pattern in both datasets is similar, suggesting prevalence in cloudy
skies over Thessaloniki.
For clear skies (Figure\ \ref{fig:cusumnotrendmonthly-2}) the distinct change
in 2004 is now clearer. The most likely reason for this change is the
monotonic reduction of aerosols in Thessaloniki. At that year there a
change in the rate of decrease in aerosol optical depth as illustrated
in Figure 7 of @Siomos2020. This abrupt change in CUSUMs lasts
until about 2010 when the anomalies become again variable.

@@

## Effects of the solar zenith angle on SDR

>>>>

Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix) shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
For time series with a uniform trend, we would expect
the CUSUMs of the anomalies to have a symmetric 'V' shape centered
around the middle of the series. This would indicate that the anomalies
are evenly distributed around the climatological mean, and, for a
positive uniform trend, the first half is below and the second half
above the climatological mean.
In our case, there is a more complex
evolution of the anomalies.
For all skies (Figure\ \ref{fig:cusummonth-1}),
we observe three
rather distinct periods: (a) a downward part between the start of the
datasets and about 2000, denoting that all anomalies are negative, thus
below the climatology; (b) a relatively steady part lasting for almost 20
years suggesting little variability in SDR anomalies; and (c) a steep
upward part to the present, indicating anomalies above the climatology. The
CUSUMs for cloudy skies (Figure\ \ref{fig:cusummonth-3}) show a similar behavior with some short-term
differences that do not change the overall pattern.
For clear skies
(Figure\ \ref{fig:cusummonth-2}), a monotonic downward tendency is evident until 2004,
suggesting that the anomalies are all negative. After 2004, the anomalies
turn positive at a fast rate for about five years and at a slower
rate thereafter.

@@

In order to unveil further the features of the variability of the three
datasets, Figure\ \ref{fig:cusumnotrendmonthly}
presents another set of CUSUM plots using anomalies
after the long-term linear trend is removed. With this approach, periods
when the CUSUMs diverge from zero can be interpreted as a systematic
variation of SDR from the climatological mean. When the CUSUM is
increasing, the anomalies values are above the climatology and vice
versa. 
Overall, for all- and cloudy-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-3}) we
observe periods with anomalies diverging from the climatological values,
each lasting for several years. These fluctuations are probably within
the natural variability and no distinct changes are identified. The
pattern in both datasets is similar, suggesting prevalence in cloudy
skies over Thessaloniki.
For clear skies (Figure\ \ref{fig:cusumnotrendmonthly-2}) the distinct change
in 2004 is now clearer. The most likely reason for this change is the
monotonic reduction of aerosols in Thessaloniki. In that year, there was a
change in the rate of decrease in aerosol optical depth, as illustrated
in Figure 7 of @Siomos2020.
This abrupt change in CUSUMs lasted
until about 2010 when the anomalies become again variable.

@@

## Effects of the Solar Zenith Angle on SDR

### 2023-12-28 MDPI_submission.tex
Commit 2023-12-28 19:09

Although the year-to-year variability of the anomalies (Figure
\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR} and
\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior,
plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
can reveal different structures in the records of all three sky
conditions. For time series with a uniform trend, we would expect the
CUSUMs of the anomalies to have a symmetric `V' shape centered around
the middle of the series. This would indicate that the anomalies are
evenly distributed around the climatological mean, and for a positive
uniform trend, the first half is below and the second half above the
climatological mean. In our case, there is a more complex evolution of
the anomalies. For all skies (Figure~\ref{fig:cusummonth-1}), we observe
three rather distinct periods: a) a downward part between the start of
the datasets and about 2000, denoting that all anomalies are negative,
thus below the climatology; b) a relatively steady part lasting for
almost 20 years suggesting little variability in SDR anomalies; and c) a
steep upward part to the present indicating anomalies above the
climatology. The CUSUMs for cloudy-skies
(Figure~\ref{fig:cusummonth-3}), show a similar behavior with some
short-term differences that do not change the overall pattern. For clear
skies (Figure~\ref{fig:cusummonth-2}), a monotonic downward tendency is
evident until 2004, suggesting that the anomalies are all negative.
After 2004 the anomalies turn to positive at a fast rate for about five
years and at a slower rate thereafter.

@@

In order to unveil further the features of the variability of the three
datasets, Figure~\ref{fig:cusumnotrendmonthly} presents another set of
CUSUM plots using anomalies after the long-term linear trend is removed.
With this approach, periods when the CUSUMs diverge from zero can be
interpreted as a systematic variation of SDR from the climatological
mean. When the CUSUM is increasing, the anomalies values are above the
climatology and vice versa. Overall, for all- and cloudy-sky conditions
(Figures~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-3}) we observe periods with anomalies
diverging from the climatological values, each lasting for several
years. These fluctuations are probably within the natural variability
and no distinct changes are identified. The pattern in both datasets is
similar, suggesting prevalence in cloudy skies over Thessaloniki. For
clear skies (Figure~\ref{fig:cusumnotrendmonthly-2}) the distinct change
in 2004 is now clearer. The most likely reason for this change is the
monotonic reduction of aerosols in Thessaloniki. At that year there a
change in the rate of decrease in aerosol optical depth as illustrated
in Figure 7 of \citet{Siomos2020}. This abrupt change in CUSUMs lasts
until about 2010 when the anomalies become again variable.

>>>>

Although the year-to-year variability of the anomalies (Figure
\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR} and
\ref{fig:trendCLOUD} in Appendix) shows a rather homogeneous behavior,
plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
can reveal different structures in the records of all three sky
conditions. For time series with a uniform trend, we would expect the
CUSUMs of the anomalies to have a symmetric `V' shape centered around
the middle of the series. This would indicate that the anomalies are
evenly distributed around the climatological mean, and, for a positive
uniform trend, the first half is below and the second half above the
climatological mean. In our case, there is a more complex evolution of
the anomalies. For all skies (Figure~\ref{fig:cusummonth-1}), we observe
three rather distinct periods: (a) a downward part between the start of
the datasets and about 2000, denoting that all anomalies are negative,
thus below the climatology; (b) a relatively steady part lasting for
almost 20 years suggesting little variability in SDR anomalies; and (c)
a steep upward part to the present, indicating anomalies above the
climatology. The CUSUMs for cloudy skies (Figure~\ref{fig:cusummonth-3})
show a similar behavior with some short-term differences that do not
change the overall pattern. For clear skies
(Figure~\ref{fig:cusummonth-2}), a monotonic downward tendency is
evident until 2004, suggesting that the anomalies are all negative.
After 2004, the anomalies turn positive at a fast rate for about five
years and at a slower rate thereafter.

@@

In order to unveil further the features of the variability of the three
datasets, Figure~\ref{fig:cusumnotrendmonthly} presents another set of
CUSUM plots using anomalies after the long-term linear trend is removed.
With this approach, periods when the CUSUMs diverge from zero can be
interpreted as a systematic variation of SDR from the climatological
mean. When the CUSUM is increasing, the anomalies values are above the
climatology and vice versa. Overall, for all- and cloudy-sky conditions
(Figures~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-3}) we observe periods with anomalies
diverging from the climatological values, each lasting for several
years. These fluctuations are probably within the natural variability
and no distinct changes are identified. The pattern in both datasets is
similar, suggesting prevalence in cloudy skies over Thessaloniki. For
clear skies (Figure~\ref{fig:cusumnotrendmonthly-2}) the distinct change
in 2004 is now clearer. The most likely reason for this change is the
monotonic reduction of aerosols in Thessaloniki. In that year, there was
a change in the rate of decrease in aerosol optical depth, as
illustrated in Figure 7 of \citet{Siomos2020}. This abrupt change in
CUSUMs lasted until about 2010 when the anomalies become again variable.
